---
title: "Reef Fish Abundance Across Time"
output: html_document
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
library(tidyverse)
reef_fish <- read.csv("Data/Supplemental201.csv")
```

## Modifying Raw Data

To begin analyzing the raw data, I used dplyr commands to select columns relevant to my research interest. I used the filter commmand to filter out rows where species were not present at either location, and to eliminate a row provided in the raw data totaling the amount of species seen at each location.

```{r}
reef_fish_bootless <- select(reef_fish, Genus, species, Bootless, Museum.Bootless, Fished) %>% 
  filter(Bootless > 0 | Museum.Bootless > 0) %>% 
  filter(Bootless < 50)
```

```{r}
reef_fish_bootless_sp <- group_by(reef_fish_bootless, Genus) %>%
  summarise(
    bootless = sum(Bootless),
    Museum.Bootless = sum(Museum.Bootless),
    n_species = n())
knitr::kable(head(reef_fish_bootless_sp, n = 20), format = "html")
```

```{r, echo=FALSE}
ggplot(reef_fish_bootless_sp) +
  geom_point(aes(x = bootless, y = n_species), color = "blue", size = 3, alpha = 0.4) +
  geom_point(aes(x = Museum.Bootless, y = n_species), color = "red", size = 3, alpha = 0.4) +
  labs(x = "Species Present by Location", y = "Total Species Count")
```



```{r, echo=FALSE}

```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
